---
title: "SAGs' global abundance"
date: "27.11.2016"
output:  
  html_document:
    theme: united
    toc: yes
  pdf_document:
    highlight: zenburn
    toc: yes
---

<!--- INITIALIZATION
```{r, echo=FALSE}
#error hook to kill knitr in case of errors
library(knitr)
knit_hooks$set(error = function(x, options) stop(x))
opts_chunk$set(cache=TRUE, autodep=TRUE)
```
--->

# 1) Overview

```{r libraries, echo=F, message=F}
library(data.table)
library(seqinr)
library(reshape2)
library(ggplot2)
library(ggrepel)
library(plyr)
library(DT)
library(shiny)
```


```{r read_V9_from_SAGs, echo=F, message=F, results="hide"}
setwd("~/Documents/2Q16/analyses/exploratory_figures/dominant_mixotrophs/")

sag_18S<-fread("SAG_18S_sequences.csv")
nrow(sag_18S) #903 SAGs

primer1<-"TTGTACACACCGCCC"

sag_18S[,V9:=grepl(primer1,Sequences)]
sag_18S<-sag_18S[V9==T]
sag_18S[,Sequences:=sub(paste("^.+",primer1,sep=""),"",Sequences)]
sag_18S[,V9:=NULL]
nrow(sag_18S) #868 SAGs
head(sag_18S)

write.fasta(sequences = as.list(sag_18S$Sequences), names=sag_18S$ID_seq,file.out = "sag_V9.fas")
```

A total amount of 868 SAGs (out of 903) contain the sequence of the forward primer 1389F for the V9 region (5’- TTGTACACACCGCCC -3’). None of the 903 SAGs contains the complete sequence of the reverse primer 1510R (5’- CCTTCYGCAGGTTCACCTAC -3’) - most of the 18S rDNA SAGs miss the last ~5 bp of the V9 region.

We blasted the obtained V9 sequences on Tara swarms and we selected the hits showing a coverage >80% and a similarity =>97%. 

```{r blast_SAGs-V9_on_swarms, echo=F, message=F, results="hide"}

#system("vsearch --usearch_global sag_V9.fas --maxrejects 0 --maxaccepts 0 --top_hits_only -db /home0/data/Tara/last/tara_ref.fas --blast6out output.txt --id 0.8")
 
output<-fread("output.txt")
output[,coverage:=V4/V10*100]
output<-output[coverage>80]
output<-output[,list(V1,V2,V3,V4,V10,coverage)]
setnames(output,c("SAG_id","md5sum","identity","SAG_length","swarm_length","coverage"))
output<-output[identity>=97]
```


```{r samples_selection, echo=F, message=F, results="hide"}
#global occurrence of swarms in Tara Oceans, BioMarKs and Tara Arctic.

#selection of Tara samples.
sple<-fread("sple.csv")
sple<-sple[fraction%in%c("0.8-5","5-20","20-180","180-2000")&depth%in%c("SUR","DCM")&template=="DNA"]

data_abund<-fread("data_abund_table.txt")
#write.table(data_abund, "data_abund_table.txt", sep="\t",row.names=T)

swarms_tb_reduced<-fread("data_info_table.txt")
head(swarms_tb_reduced)
#write.table(data[,.SD,.SDcols=colnames(data)[!grepl("TV|TA|BV",colnames(data))]],"data_info_table.txt",sep="\t",row.names=T)

#temp<-data[taxogroup=="Prasinophyceae Clade 7"]
#rbind.fill.matrix(lapply(strsplit(temp$lineage,"\\|"),function(X) t(as.matrix(X))))
```


```{r read_SAGs_classification_and_status, echo=F, message=F, results="hide"}
SAGs_classification<-read.table(file="SAGs_complete_IDs.txt", head=TRUE, fill=TRUE)
SAGs_classification[1:50,]

SAGs_status<-read.table(file="output_classif.txt", head=TRUE, fill=TRUE)
SAGs_status<-as.data.frame(SAGs_status)
SAGs_status[1:5,]
```


```{r fig_all, echo=F, include=T, results="hide", message=F, warning=F, fig.align="center"}
#add swarms taxogroup data
#data_abund$swm_taxogroup = swarms_tb_reduced[match(data_abund$md5sum, swarms_tb_reduced$md5sum), "taxogroup"] 
#data_abund[1:5,]

temp<-data_abund[,sum(value),by=md5sum]
setkey(temp,md5sum)
setkey(output,md5sum)
temp<-output[temp,allow.cartesian=T]
temp<-temp[order(V1,decreasing = T)]
temp[,rank:=1:nrow(temp)]
temp<-temp[V1>10]
temp[,SAG:=sapply(identity,function(X) ifelse(is.na(X),"no","yes"))]
temp<-join(temp, SAGs_classification)
temp<-join(temp, SAGs_status)
temp$swm_taxogroup<-swarms_tb_reduced[match(temp$md5sum, swarms_tb_reduced$md5sum), "taxogroup"] 
temp[1:50,]

swarms_groups <- list(
  "no"="no SAG aligned",
  "yes"="SAG(s) aligned"
)

swarms_labeller <- function(variable,value){
  return(swarms_groups[value])
}

ggplot(temp,aes(x=rank,y=V1))+
  geom_point(aes(colour=SAG))+
  scale_y_log10(name="abundance")+
  facet_wrap(~SAG,ncol=2, labeller = swarms_labeller)+
  scale_colour_discrete(name="swarms",
                    breaks=c("no", "yes"),
                    labels=c("no SAG aligned", "SAG(s) aligned"))+
  #theme(legend.position="bottom")+
  labs(title="[Fig. 1] All swarms")

ggsave("SAGs_abund_fig_all.png",height=6.75,width=9)
```

No. of swarms with SAG(s) aligned:

```{r all_yes_length, echo=F}
length(which(temp$SAG == "yes")) #872
```

No. of swarms with no SAG aligned:

```{r all_no_length, echo=F}
length(which(temp$SAG == "no")) #92837
```

```{r all_table_dataset, echo=F, results="hide"}
SAGs_lt<-temp[temp$SAG =="yes"]

SAGs_lt$id_classif<-as.character(SAGs_lt$id_classif)

#round coverage values
SAGs_lt[,'coverage']<-round(SAGs_lt[,'coverage'],3)

#reorder columns
SAGs_lt<-SAGs_lt[,c(10,2,3,4,5,6,7,8,12,11)]
SAGs_lt[1:5,]
```
\
\
\
```{r all_table, echo=F}

SAGs_lt<-as.data.frame(SAGs_lt)

datatable(head(SAGs_lt, 872), 
          colnames = c('ident.%' = 4, 'SAG_lth' = 5, 'swm_lth' = 6, 'abund' = 8),
          filter = list(position = 'top', clear = FALSE, plain = TRUE),
          class = 'cell-border stripe',
          caption = 'Table 1: abundance of all SAGs',
          
          #formatStyle(
          #  'status',
          #  backgroundColor = styleEqual(
          #    unique(SAGs_lt$status), c('orange', 'lightgreen', 'lightpink')
          #  )
          #),
          
          options = list(
            
            autowidth = TRUE,
            searchHighlight = TRUE,
            
            #1rst: abund; 2nd: coverage; 3rd: id%
            order = (list(list(7, 'desc'), list(6, 'desc'), list(3, 'desc'))),
            
            #data position in columns
            columnDefs = list(list(
             # className = "dt-center", targets=c(1:10))),
            
            #swarms IDs rendering
              #list(list(
                targets = 2,
                render = JS(
                  "function(data, type, row, meta) {",
                  "return type === 'display' && data.length > 6 ?",
                  "'<span title=\"' + data + '\">' + data.substr(0, 6) + '...</span>' : data;",
                  "}"))),
          
            #head layout
            initComplete = JS(
            "function(settings, json) {",
            "$(this.api().table().header()).css({'background-color': '#323232', 'color': '#fff'});",
            "}"),
            
            pageLength = 10,
            lengthMenu = c(10, 20, 50, 100, 300, 872)
))

#SAGs_lt[1:100,]
#write.table(SAGs_lt, "SAGs_abundance_tb_all.txt", sep="\t")
#app = system.file('examples', 'DT-rows', package = 'DT')
#shiny::runApp(app)
```

\
\
\
\

SURFACE - Fraction 0.8-5

```{r fig_piconano, echo=F, results="hide"}
data_abund_SUR<-data_abund[depth=="SUR"]
temp<-data_abund_SUR[fraction=="0.8-5",sum(value),by=list(md5sum,fraction)]
setkey(temp,md5sum)
setkey(output,md5sum)
temp<-output[temp,allow.cartesian=T]
temp<-temp[order(V1,decreasing = T)]
temp[,rank:=1:nrow(temp)]
temp<-temp[V1>10]
temp[,SAG:=sapply(identity,function(X) ifelse(is.na(X),"no","yes"))]
temp<-join(temp, SAGs_classification)
temp[1:12,]

ggplot(temp,aes(x=rank,y=V1))+
  geom_point(aes(colour=SAG))+
  scale_x_continuous(limits=c(0,45000), breaks = c(0,20000,40000))+
  scale_y_log10()+
  facet_wrap(~SAG,ncol=2, labeller = swarms_labeller)+
  scale_colour_discrete(name="swarms",
                    breaks=c("no", "yes"),
                    labels=c("no SAG aligned", "SAG(s) aligned"))+
  theme(legend.position="bottom")
  geom_text_repel(data = subset(temp, temp$SAG=="yes" & temp$V1>900000), aes(x=rank,y=V1,label=id_classif), size=2.5, force=2)

ggsave("SAGs_abundance_fig_piconano_SRF_lab.png",height=6.75,width=9)


ggplot(temp,aes(x=rank,y=V1))+
  geom_point(aes(colour=SAG))+
  scale_x_continuous(limits=c(0,45000), breaks = c(0,20000,40000))+
  scale_y_log10()+
  facet_wrap(~SAG,ncol=2, labeller = swarms_labeller)+
  scale_colour_discrete(name="swarms",
                    breaks=c("no", "yes"),
                    labels=c("no SAG aligned", "SAG(s) aligned"))+
  theme(legend.position="bottom")
 
ggsave("SAGs_abundance_fig_piconano_SRF_lab.png",height=6.75,width=9)


  
# size=2, force=3, box.padding = unit(0.1,"lines"), point.padding = unit(0.02, "lines"))

ggplot(data=subset(temp, temp$SAG=="yes" & temp$V1>100000),aes(x=rank,y=V1))+
  geom_point(aes(colour=SAG))+
  scale_x_continuous(limits=c(0,45000), breaks = c(0,20000,40000))+
  scale_y_log10()+
  geom_text_repel(data = subset(temp, temp$SAG=="yes" & temp$V1>500000), aes(x=rank,y=V1,label=id_classif), size=2.5)

length(which(temp$SAG == "yes")) #817
length(which(temp$SAG == "no")) #38465

SAGs_lt<-temp[temp$SAG =="yes"]
SAGs_lt[1:50,]
```

Fraction 5-20um

```{r fig_nano, echo=F}
temp<-data_abund[fraction=="5-20",sum(value),by=list(md5sum,fraction)]
setkey(temp,md5sum)
setkey(output,md5sum)
temp<-output[temp,allow.cartesian=T]
temp<-temp[order(V1,decreasing = T)]
temp[,rank:=1:nrow(temp)]
temp<-temp[V1>10]
temp[,SAG:=sapply(identity,function(X) ifelse(is.na(X),"no","yes"))]

ggplot(temp,aes(x=rank,y=V1))+geom_point(aes(colour=SAG))+scale_y_log10()+facet_wrap(~SAG,ncol=2)

length(which(temp$SAG == "yes")) #681
length(which(temp$SAG == "no")) #28134

SAGs_lt<-temp[temp$SAG =="yes"]
SAGs_lt[1:50,]
```

Fraction 20-180um

```{r fig_micro, echo=F}
temp<-data_abund[fraction=="20-180",sum(value),by=list(md5sum,fraction)]
setkey(temp,md5sum)
setkey(output,md5sum)
temp<-output[temp,allow.cartesian=T]
temp<-temp[order(V1,decreasing = T)]
temp[,rank:=1:nrow(temp)]
temp<-temp[V1>10]
temp[,SAG:=sapply(identity,function(X) ifelse(is.na(X),"no","yes"))]

ggplot(temp,aes(x=rank,y=V1))+geom_point(aes(colour=SAG))+scale_y_log10()+facet_wrap(~SAG,ncol=2)

length(which(temp$SAG == "yes")) #678
length(which(temp$SAG == "no")) #24988

SAGs_lt<-temp[temp$SAG =="yes"]
SAGs_lt[1:50,]
```

Fraction 180-2000um

```{r fig_meso, echo=F}
temp<-data_abund[fraction=="180-2000",sum(value),by=list(md5sum,fraction)]
setkey(temp,md5sum)
setkey(output,md5sum)
temp<-output[temp,allow.cartesian=T]
temp<-temp[order(V1,decreasing = T)]
temp[,rank:=1:nrow(temp)]
temp<-temp[V1>10]
temp[,SAG:=sapply(identity,function(X) ifelse(is.na(X),"no","yes"))]

ggplot(temp,aes(x=rank,y=V1))+geom_point(aes(colour=SAG))+scale_y_log10()+facet_wrap(~SAG,ncol=2)

length(which(temp$SAG == "yes")) #683
length(which(temp$SAG == "no")) #24775

SAGs_lt<-temp[temp$SAG =="yes"]
SAGs_lt[1:50,]
```
